%
% LaTeX template for prepartion of submissions to PLDI'16
%
% Requires temporary version of sigplanconf style file provided on
% PLDI'16 web site.
% 
\documentclass[pldi]{sigplanconf-pldi16}
% \documentclass[pldi-cameraready]{sigplanconf-pldi16}

%
% the following standard packages may be helpful, but are not required
%
\usepackage{SIunits}            % typset units correctly
\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
% known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}   % print a hyperlinked DOI



\begin{document}

\title{Instructions for Submission to PLDI'16}

%
% any author declaration will be ignored  when using 'pldi' option (for double blind review)
%

\authorinfo{Person 1 \and Person 2}
{\makebox{A Department} \\
\makebox{A University}  \\
\makebox{A Place, AS 12345}}
{\{person1,person2\}@cs.auniv.edu}

\maketitle

\begin{abstract}
  We present programming by example that can utilize user defined higher order functions.
  We use refine,ent types to prune the search space of higher order functions.
  Since our refinement types can be under approximating, we can extend liquid haskell to support arbitraty syntax extensions.
  We introduce the concept of dismantling procedures to guide first order function synthesis.
  We use results from paramatricity as our dismantling procedure.
\end{abstract}

\section{Introduction}

Here is the paper

\bibliographystyle{abbrvnat}

% We recommend that you use BibTeX.  The inlined bibitems below are
% used to keep this template to a single file.
\begin{thebibliography}{}
\softraggedright

\bibitem[Backus et~al.(1960)]{Backus:60}
J.~W. Backus, F.~L. Bauer, J.~Green, C.~Katz, J.~McCarthy, A.~J. Perlis,
  H.~Rutishauser, K.~Samelson, B.~Vauquois, J.~H. Wegstein, A.~van Wijngaarden,
  and M.~Woodger.
\newblock Report on the algorithmic language ALGOL 60.
\newblock \emph{Commun. ACM}, 3\penalty0 (5):\penalty0 299--314, May 1960.
\newblock ISSN 0001-0782.
\newblock \doi{10.1145/367236.367262}.

\bibitem[Collins(1960)]{Collins:60}
G.~E. Collins.
\newblock A method for overlapping and erasure of lists.
\newblock \emph{Commun. ACM}, 3\penalty0 (12):\penalty0 655--657, December
  1960.
\newblock \doi{10.1145/367487.367501}

\bibitem[Lamport(1994)]{Lamport:94}
L.~Lamport.
\newblock \emph{{\LaTeX: A Document Preparation System}}.
\newblock Addison-Wesley, Reading, Massachusetts, 2nd edition, 1994.

\bibitem[McCarthy(1960)]{McCarthy:60}
J.~McCarthy.
\newblock Recursive functions of symbolic expressions and their computation by
  machine, part {I}.
\newblock \emph{Commun. ACM}, 3\penalty0 (4):\penalty0 184--195, April 1960.
\newblock \doi{10.1145/367177.367199}

\end{thebibliography}

\end{document}
